Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2021-10-21 analysis/paper_draft2.Rmd
updated

And as our prototype app shows, if the historical strike data loaded into the app in advance, attorneys and others can the approach here in real time during jury selection. Despite _Batson_ and similar law, racial bias persists in the use of peremptory challenges [@Diamond2018], perhaps in part because attorneys fear that _raising_ a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021]. It is worth studying whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges.

@@

Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data is, for each attorney, a complete history of who they struck in all their past cases. The problem: Jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018]. For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender). This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information. And judges and court staff vary in how often they fully complete or maintain the information they are required to record. Incomplete historical strike data resents an issue for inferring bias if such data is incomplete or missing not at random. These data challenges, however, might decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.

>>>>

Moreover, if the historical strike data can loaded into a tool like our prototype app, attorneys and others can the approach here in real time during jury selection. Despite _Batson_ and similar law, racial disparity persists in the use of peremptory challenges [@Diamond2018]. One concern is that attorneys fear that arguing _Batson_  will produce the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021]. At the same time, we assume that lawyers and judges want to discourage an attorney from raising a _Batson_ challenge without a credible basis. It is worth studying whether an app like the prototype presented here might reduce such barriers to bringing otherwise credible _Batson_ challenges.

@@

Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data is, for each attorney, a complete history of who they struck in all their past cases. The problem: Jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018]. For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender). This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information. And judges and court staff vary in how often they fully complete or maintain the information they are required to record. Incomplete historical strike data presents an issue for inferring bias if such data is incomplete or missing not at random. These data challenges, however, might decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
